---
title: "Women's Soccer Data"
output: html_notebook
---

We'll use this notebook to explore some NCAA women's soccer data and [our questions](https://docs.google.com/spreadsheets/d/1kTSrnl1URmcexIWLDMWqDnp3q-c2OS45uifnhjp10n4/edit#gid=0). First, let's load the tidyverse library:

```{r}
library(tidyverse)
```

We'll want to start with 2022 match data, which is located in [this GitHub repository](https://github.com/dwillis/NCAAWomensSoccerData). We can load the match data directly from GitHub using `read_csv`:


```{r}
matches_2022 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2022.csv")
```

Let's take a look at this data to see what it contains, using `head` and `glimpse`:

```{r}
head(matches_2022)
glimpse(matches_2022)
```

Now, onto the questions using aggregates, which means `group_by` and `summarize`:

1. How many shots in total has Maryland attempted and how many have been on goal? Calculate a percentage of shots on goal.

```{r}
matches_2022 %>% 
 filter(team=="Maryland Terrapins, Terps") %>% 
 group_by(team) %>% 
 summarize(total_shots = sum(sh_att), qual_shot = sum(so_g)) %>% 
 mutate(per_qual = (qual_shot/total_shots)*100)

```

2. What about the same shot stats from last season's first five games? You'll need to load the matches from 2021 - just change the season.

```{r}
matches_2021 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2021.csv")

matches_2021 %>% 
 filter(team=="Maryland Terrapins, Terps", date < "2021-09-05") %>% 
  group_by(team) %>% 
 summarize(total_shots = sum(sh_att), qual_shot = sum(so_g)) %>% 
 mutate(per_qual = (qual_shot/total_shots)*100)
  
 
```


3. What about draws - Maryland started with 4 ties. Have any other teams had as many?
NJIT, Pacific, St Bonaventure and VCU are the other teams with as many draws.
```{r}
matches_2022 %>% 
  filter(result =="T") %>% 
  group_by(team) %>% 
  summarize(count = n()) %>% 
  filter(count >= 4)
```


4. How many total corner kicks has Maryland had so far and how many have its opponents have? Calculate the difference between the two totals for every team.

Maryland has 31 corners for and 24 against, giving them a difference of seven.
```{r}
matches_2022 %>% 
  filter(team == "Maryland Terrapins, Terps") %>% 
  group_by(team) %>% 
  summarize(offcorner = sum(corners), defcorner = sum(defensive_corners), difference = offcorner-defcorner)

matches_2022 %>% 
  group_by(team) %>% 
  summarize(offcorner = sum(corners), defcorner = sum(defensive_corners), difference = offcorner-defcorner) %>% 
  arrange(desc(difference))
```
